### 2013-10-07 abstract.tex
updated text to reflect new volume data (Fig 4)

\section*{Abstract}
Prostate cancer (PCa) is the most common non-cutaneous malignancy among men in
the United States and the second leading cause of cancer related death.
Non-invasive imaging could lead to improved diagnosis, risk-stratification, and
PCa management.  Magnetic resonance imaging (MRI) has been available for use in
the workup of patients with PCa since the early 1980s, and recent advances with
functional parameters has greatly improved its clinical diagnostic utility.
Acoustic Radiation Force Impulse (ARFI) imaging is an ultrasound-based modality
that evaluates the mechanical properties of soft tissues. ARFI imaging has the
potential to aid in PCa diagnosis and management by evaluating the structural
composition of prostate zones and tumors based on their stiffness.  In this
study, MR and ARFI imaging datasets were compared to one another and with gross
pathology measurements made immediately post radical prostatectomy.  Imaging
datasets were manually segmented to delineate the central gland and prostate
capsule, and 3D models were rendered to evaluate zonal anatomy dimensions and
volumes.  Both imaging modalities showed moderate correlations (0.39 $<$ R$^2 <
$ 0.74) between estimated organ volume and gross pathologic weights.  ARFI and MR
total prostate gland volumes were well-correlated (R$^2$ = 0.68), but ARFI
images yielded prostate volumes that were, on average, larger (36\% $\pm$ 28\%)
than MR images, primarily due to over-estimation of the anterior-to-posterior
dimension of the prostate total gland (17.0 $\pm$ 12.1\%), while over-estimates
of the other dimensions were less significant contributors (8.1 $\pm$ 18.4\%
and 0.58 $\pm$ 12.9\%).  The central zone volumes of ARFI and MR images were
also moderately correlated (R$^2$ = 0.41), with minimal volume bias between the
imaging modalities, but significant variability case-to-case (2.1 $\pm$
39.1\%).  Central zone volume differences were, again, strongly attributed to
over-estimation of the anterior-to-posterior axis (14.8 $\pm$ 23.1\%), with a
significant underestimation of the apex-to-base dimension (-10.8 $\pm$ 22.3\%)
and no mean bias in the lateral-to-lateral measurements (0.006 $\pm$ 17.2\%).
Strong variability in central gland volumes is believed to be related to the
extent of benign prostatic hyperplasia (BPH) for select cases.  Overall, ARFI
imaging of the prostate yielded prostate volumes and dimensions that were
correlated with MR T2WI estimates, with biases in the anterior-to-posterior
dimension, most likely related to poor displacement SNR in the anterior region
of the prostate from greater distance from the rectal wall imaging surface.
ARFI imaging is a promising low-cost, real-time imaging modality that can
compliment MR imaging for diagnosis, treatment planning and management of PCa.

>>>>

\section*{Abstract}
Prostate cancer (PCa) is the most common non-cutaneous malignancy among men in
the United States and the second leading cause of cancer related death.
Non-invasive imaging could lead to improved diagnosis, risk-stratification, and
PCa management.  Magnetic resonance imaging (MRI) has been available for use in
the workup of patients with PCa since the early 1980s, and recent advances with
functional parameters has greatly improved its clinical diagnostic utility.
Acoustic Radiation Force Impulse (ARFI) imaging is an ultrasound-based modality
that evaluates the mechanical properties of soft tissues. ARFI imaging has the
potential to aid in PCa diagnosis and management by evaluating the structural
composition of prostate zones and tumors based on their stiffness.  In this
study, MR and ARFI imaging datasets were compared to one another and with gross
pathology measurements made immediately post radical prostatectomy.  Imaging
datasets were manually segmented to delineate the central gland and prostate
capsule, and 3D models were rendered to evaluate zonal anatomy dimensions and
volumes.  Both imaging modalities showed moderate correlations (0.39 $<$ R$^2 <
$ 0.74) between estimated organ volume and gross pathologic weights.  ARFI and MR
total prostate gland volumes were well-correlated (R$^2$ = 0.63), but ARFI
images yielded prostate volumes that were, on average, larger (6.1\% $\pm$ 25\%)
than MR images, primarily due to over-estimation of the anterior-to-posterior
dimension of the prostate total gland (17.0 $\pm$ 12.1\%), while over-estimates
of the other dimensions were less significant contributors (8.1 $\pm$ 18.4\%
and 0.58 $\pm$ 12.9\%).  The central zone volumes of ARFI and MR images were
also moderately correlated (R$^2$ = 0.38), with minimal volume bias between the
imaging modalities, but significant variability case-to-case (-5.0 $\pm$
39.5\%).  Central zone volume differences were, again, strongly attributed to
over-estimation of the anterior-to-posterior axis (14.8 $\pm$ 23.1\%), with a
significant underestimation of the apex-to-base dimension (-10.8 $\pm$ 22.3\%)
and no mean bias in the lateral-to-lateral measurements (0.006 $\pm$ 17.2\%).
Strong variability in central gland volumes is believed to be related to the
extent of benign prostatic hyperplasia (BPH) for select cases.  Overall, ARFI
imaging of the prostate yielded prostate volumes and dimensions that were
correlated with MR T2WI estimates, with biases in the anterior-to-posterior
dimension, most likely related to poor displacement SNR in the anterior region
of the prostate from greater distance from the rectal wall imaging surface.
ARFI imaging is a promising low-cost, real-time imaging modality that can
compliment MR imaging for diagnosis, treatment planning and management of PCa.

### 2013-10-07 conclusions.tex
updated text to reflect new volume data (Fig 4)

\section{Conclusions}
The delineation of prostate zonal anatomy in ARFI images has been compared with
the established methods for identifying zonal anatomy using MR T2W images.
Both imaging modalities showed moderate correlations (0.39 $<$ R$^2 < $ 0.74)
between estimated organ volume and gross pathologic weights.  ARFI and MR total
prostate gland volumes were well-correlated (R$^2$ = 0.68), but ARFI images
yielded prostate volumes that were, on average, larger (36\% $\pm$ 28\%) than
MR images, primarily due to over-estimation of the anterior-to-posterior
dimension of the prostate total gland (17.0 $\pm$ 12.1\%), while over-estimates
of the other dimensions were less significant contributors (8.1 $\pm$ 18.4\%
and 0.58 $\pm$ 12.9\%).  The central zone volumes of ARFI and MR images were
also moderately correlated (R$^2$ = 0.41), with minimal volume bias between the
imaging modalities, but significant variability case-to-case (2.1 $\pm$
39.1\%).  Central zone volume differences were, again, strongly attributed to
over-estimation of the anterior-to-posterior axis (14.8 $\pm$ 23.1\%), with a
significant underestimation of the apex-to-base dimension (-10.8 $\pm$ 22.3\%)
and no mean bias in the lateral-to-lateral measurements (0.006 $\pm$ 17.2\%).
Strong variability in central gland volumes is believed to be related to the
extent of benign prostatic hyperplasia (BPH) for select cases.  Overall, ARFI
imaging of the prostate yielded prostate volumes and dimensions that were
correlated with MR T2WI estimates, with biases in the anterior-to-posterior
dimension, most likely related to poor displacement SNR in the anterior region
of the prostate from greater distance from the rectal wall imaging surface.
ARFI imaging is a promising low-cost, real-time imaging modality that can
compliment MR imaging for diagnosis, treatment planning and management of PCa.

>>>>

\section{Conclusions}
The delineation of prostate zonal anatomy in ARFI images has been compared with
the established methods for identifying zonal anatomy using MR T2W images.
Both imaging modalities showed moderate correlations (0.39 $<$ R$^2 < $ 0.74)
between estimated organ volume and gross pathologic weights.  ARFI and MR total
prostate gland volumes were well-correlated (R$^2$ = 0.63), but ARFI images
yielded prostate volumes that were, on average, larger (6.1\% $\pm$ 25\%) than
MR images, primarily due to over-estimation of the anterior-to-posterior
dimension of the prostate total gland (17.0 $\pm$ 12.1\%), while over-estimates
of the other dimensions were less significant contributors (8.1 $\pm$ 18.4\%
and 0.58 $\pm$ 12.9\%).  The central zone volumes of ARFI and MR images were
also moderately correlated (R$^2$ = 0.38), with minimal volume bias between the
imaging modalities, but significant variability case-to-case (-5.0 $\pm$
39.5\%).  Central zone volume differences were, again, strongly attributed to
over-estimation of the anterior-to-posterior axis (14.8 $\pm$ 23.1\%), with a
significant underestimation of the apex-to-base dimension (-10.8 $\pm$ 22.3\%)
and no mean bias in the lateral-to-lateral measurements (0.006 $\pm$ 17.2\%).
Strong variability in central gland volumes is believed to be related to the
extent of benign prostatic hyperplasia (BPH) for select cases.  Overall, ARFI
imaging of the prostate yielded prostate volumes and dimensions that were
correlated with MR T2WI estimates, with biases in the anterior-to-posterior
dimension, most likely related to poor displacement SNR in the anterior region
of the prostate from greater distance from the rectal wall imaging surface.
ARFI imaging is a promising low-cost, real-time imaging modality that can
compliment MR imaging for diagnosis, treatment planning and management of PCa.

### 2013-10-07 fig_mr_arfi_volumes.tex
updated text to reflect new volume data (Fig 4)

\begin{figure}[htb!]
\centering
\begin{tabular}{ccc}
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_volumes} &
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_total_linreg} &
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_central_linreg} \\
(a) CG:PZ Volume & (b) Total Gland Volume & (c) Central Gland Volume \\
\end{tabular}
\caption{Comparison of MR and ARFI zonal anatomy volume estimates from
    manually-segmented images.  Total prostate gland volumes ranged from 19.6--71.0
    cm$^3$ based on MR image models (a), with a moderate correlation between MR
    and ARFI imaging for both the total gland volume (R$^2$ = 0.68, (b)) and the
    central gland (R$^2$ = 0.41, (c)).  The ARFI total gland volumes had a positive bias
    compared with the MR volumes (36.7 $\pm$ 27.9\% (b)), while the ARFI
    central zone volumes were on average more similar (2.1 $\pm$ 39.0\%, (c)), but
    with some significant variability between study subjects.
    Table~\ref{tab:mr_arfi_volumes} contains the individual volume estimates
    for the total prostate and central glands.}
\label{fig:mr_arfi_volumes} 
\end{figure}

>>>>

\begin{figure}[htb!]
\centering
\begin{tabular}{ccc}
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_volumes} &
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_total_linreg} &
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_central_linreg} \\
(a) CG:PZ Volume & (b) Total Gland Volume & (c) Central Gland Volume \\
\end{tabular}
\caption{Comparison of MR and ARFI zonal anatomy volume estimates from
    manually-segmented images.  Total prostate gland volumes ranged from
    19.6--71.0 cm$^3$ based on MR image models (a), with a moderate correlation
    between MR and ARFI imaging for both the total gland volume (R$^2$ = 0.63,
    (b)) and the central gland (R$^2$ = 0.38, (c)).  The ARFI total gland
    volumes had a positive bias compared with the MR volumes (6.1 $\pm$ 25.0\%
    (b)), while the ARFI central zone volumes were on average more similar
    (-5.0 $\pm$ 39.5\%, (c)), but with some significant variability between
    study subjects.  Table~\ref{tab:mr_arfi_volumes} contains the individual
    volume estimates for the total prostate and central glands.}
\label{fig:mr_arfi_volumes} 
\end{figure}

### 2013-10-07 results.tex
updated text to reflect new volume data (Fig 4)

\section{Results}\label{sect:results}
The subvolumes associated with the zonal anatomy in each imaging modality were
measured (Figure~\ref{fig:mr_arfi_volumes}(a)), with moderate correlation
between the ARFI and MR total prostate gland volumes (R$^2$ = 0.68), with a
mean overestimation of of 36.7 $\pm$ 27.9\% by ARFI imaging compared to MR
volumes (Figure~\ref{fig:mr_arfi_volumes}(b)).  Central gland volumes were
slightly less correlated between ARFI and MR images (R$^2$ = 0.41) with no
significant mean over/under-estimation (2.1 $\pm$ 39.1\%), but significant
variability between the cases (Figure~\ref{fig:mr_arfi_volumes}(c)).
Table~\ref{tab:mr_arfi_volumes} has the individually-measured volumes for MR
and ARFI imaging for each study subject.

>>>>

\section{Results}\label{sect:results}
The subvolumes associated with the zonal anatomy in each imaging modality were
measured (Figure~\ref{fig:mr_arfi_volumes}(a)), with moderate correlation
between the ARFI and MR total prostate gland volumes (R$^2$ = 0.63), with a
mean overestimation of of 6.1 $\pm$ 25.0\% by ARFI imaging compared to MR
volumes (Figure~\ref{fig:mr_arfi_volumes}(b)).  Central gland volumes were
slightly less correlated between ARFI and MR images (R$^2$ = 0.38) with no
significant mean over/under-estimation (-5.0 $\pm$ 39.5\%), but significant
variability between the cases (Figure~\ref{fig:mr_arfi_volumes}(c)).
Table~\ref{tab:mr_arfi_volumes} has the individually-measured volumes for MR
and ARFI imaging for each study subject.
